Paper Analysis

### 2025-04-11 Paper_WriteUp/Paper_Plots.qmd
Revised figure 1 to a single figure with 3 Gms and 3 1700-1900s. Highlighting Magnus and Alireza

## to load some plot defaults
source(here("analysis", "helper.R"))
source(here("analysis", "plot_templates.R"))

@@

```{r figure 1}
## Sydney's code for this figure

player_cohort_1 <- "sandro2502"

path_id <- 1  ## for 1700-1900 
data_path <- all_data_path[path_id]
files <- list.files(data_path)
lichess_data <- files |> 
  map_dfr(~read_player(data_path, .x))

bullet_data_fig_1 <- lichess_data |> 
  mutate(Event = tolower(Event)) |>
  filter(Variant == "Standard") |>
  filter(grepl("rated bullet game", Event))
rm(lichess_data)

p1 <- bullet_data |> 
  filter(Username == player_cohort_1)

p1$newELO <- ifelse(p1$White == paste0(player_cohort_1),
                    p1$WhiteElo, p1$BlackElo)
# p1$datetime <- as.POSIXct(paste(p1$UTCDate, p1$UTCTime),
#                           format = "%Y.%m.%d %H:%M:%S")
ltypes <- c(1, 2, 3, 1, 2, 3, 1, 2, 3, 1, 2, 3)
thetitle <- paste0("Player From 1700-1900 Cohort")
linewd <- 1
ylim1 <- 1200
ylim2 <- 3500
ndays <- 1

p1_sorted <- p1 |> arrange(UTCDate, UTCTime)
p1_sorted$event <- 1:nrow(p1_sorted)

tidy_plot1 <- ggplot(p1_sorted, aes(x = event, y = newELO)) +
  # scale_colour_manual(values = getPalette(colourCount)) +
  scale_linetype_manual(values = ltypes) +
  xlab("Games Played") +
  ylab("Lichess Rating") +
  ggtitle(thetitle) +
  ylim(ylim1, ylim2) +
  geom_line(lwd = linewd) +
  theme(
    plot.title = element_text(hjust = 0, vjust = 1, size = rel(1.7)),
    panel.background = element_blank(),
    plot.background = element_blank(),
    panel.grid.major.y = element_line(color = "gray75", linetype = "dotted"),
    panel.grid.major.x = element_blank(),
    panel.grid.minor = element_blank(),
    strip.background = element_blank(),
    strip.text = element_text(size = rel(1.1)),
    text = element_text(color = "gray20", size = 10),
    axis.text = element_text(size = rel(1.0)),
    axis.text.x = element_text(color = "gray20", size = rel(1.0)),
    axis.text.y = element_text(color = "gray20", size = rel(1.0)),
    axis.title.x = element_text(size = rel(1.0), vjust = 0),
    axis.title.y = element_text(size = rel(1.0), vjust = 1),
    axis.ticks.y = element_blank(),
    axis.ticks.x = element_blank(),
    legend.position = "none"
  )

player_cohort_gm <- "alireza2003"

## load in the GM data for this
path_id <- 4  ## for GM 
data_path <- all_data_path[path_id]
files <- list.files(data_path)
lichess_data <- files |> 
  map_dfr(~read_player(data_path, .x))

bullet_data_fig_1_gm <- lichess_data |> 
  mutate(Event = tolower(Event)) |>
  filter(Variant == "Standard") |>
  filter(grepl("rated bullet game", Event))
rm(lichess_data)

p2 <- bullet_data_fig_1_gm |> 
  filter(Username == player_cohort_gm)

p2$newELO <- ifelse(p2$White == paste0(player_cohort_gm),
                    p2$WhiteElo, p2$BlackElo)
thetitle <- paste0("Player From GM Cohort")
p2_sorted <- p2 |> arrange(UTCDate, UTCTime)
p2_sorted$event <- 1:nrow(p2_sorted)

tidy_plot2 <- ggplot(p2_sorted, aes(x = event, y = newELO)) +
  # scale_colour_manual(values = getPalette(colourCount)) +
  scale_linetype_manual(values = ltypes) +
  xlab("Games Played") +
  ylab("Lichess Rating") +
  ggtitle(thetitle) +
  ylim(ylim1, ylim2) +
  geom_line(lwd = linewd) +
  theme(
    plot.title = element_text(hjust = 0, vjust = 1, size = rel(1.7)),
    panel.background = element_blank(),
    plot.background = element_blank(),
    panel.grid.major.y = element_line(color = "gray75", linetype = "dotted"),
    panel.grid.major.x = element_blank(),
    panel.grid.minor = element_blank(),
    strip.background = element_blank(),
    strip.text = element_text(size = rel(1.1)),
    text = element_text(color = "gray20", size = 10),
    axis.text = element_text(size = rel(1.0)),
    axis.text.x = element_text(color = "gray20", size = rel(1.0)),
    axis.text.y = element_text(color = "gray20", size = rel(1.0)),
    axis.title.x = element_text(size = rel(1.0), vjust = 0),
    axis.title.y = element_text(size = rel(1.0), vjust = 1),
    # axis.ticks.y = element_blank(),
    axis.ticks.x = element_blank(),
    legend.position = "none"
  )

tdy1 <- tidy_plot1 + theme_single_y()
tdy2 <- tidy_plot2 + theme_single_y()

tidy_plot <- ggarrange(tdy1, tdy2)

ggsave(paste0(newfilename, ".png"), tidy_plot,
  dpi = 600, height = 5, width = 12
)
ggsave(plot = tidy_plot, 
       filename = paste0(here("Paper_WriteUp", "paper_figures",
                              "elo_plot_2_players.png")), dpi = 600,
       height = 5, width = 12)
```

>>>>

## to load some plot defaults
source(here("utils", "helper.R"))
source(here("utils", "plot_templates.R"))

@@

```{r figure 1}
#Adam - Revised Apr 11, 2025
#Updated to a single figure with 3 GMs and 3 1700-1900
#Highlighting Magnus and Alireza

#import data from 1700-1900 and gm cohorts
data_path_gm = paste0(all_data_path[4], "/")
data_path_17_19 = paste0(all_data_path[1], "/")
files_gm = list.files(data_path_gm, pattern = ".csv")
files_17_19 = list.files(data_path_17_19, pattern = ".csv")

lichess_data_gm = files_gm %>%
  map_dfr(~read_player(data_path_gm, .x))
lichess_data_17_19 = files_17_19 %>%
  map_dfr(~read_player(data_path_17_19, .x))

#combine cohorts
lichess_data = rbind(lichess_data_17_19, lichess_data_gm)

#selecting 3 users from each cohort that have around 10,000 games
users = c("larrywheels", "SeanBambic", "dewang_007", #1700-1900s
          "DrNykterstein", "alireza2003", "BlueGreensun") #gms

#data for plot
plot_data = lichess_data %>%
  mutate(Event = tolower(Event)) %>%
  filter(Event == "rated bullet game", Variant == "Standard", Username %in% users) %>%
  mutate(focal_rating = ifelse(Username == White, WhiteElo, BlackElo)) %>%
  group_by(Username) %>%
  arrange(UTCDate, UTCTime, .by_group = TRUE) %>%
  mutate(game = row_number(),
         Username = case_when( #rename magnus and alireza usernames
           Username == "DrNykterstein" ~ "DrNykterstein (Magnus Carlsen)",
           Username == "alireza2003" ~ "alireza2003 (Alireza Firouzja)",
           .default = Username
         ),
         Username = factor(Username, levels = c("DrNykterstein (Magnus Carlsen)", "alireza2003 (Alireza Firouzja)", 
                                                "BlueGreensun", "SeanBambic", "dewang_007", "larrywheels"))) %>% #for order in legend
  ungroup() %>%
  select(Username, game, focal_rating)

#plotting
rating_plot = ggplot(data = plot_data, mapping = aes(x = game, y = focal_rating, colour = Username)) +
  geom_line() +
  labs(x = "Games Played", y = "Rating") + 
  theme_single_y_legend()
rating_plot

#save
ggsave(plot = rating_plot, 
       filename = here("Paper_WriteUp", "paper_figures", "fig1.png"), dpi = 1000,
       height = 4.5, width = 10)
```
